---
title: "Baseline data simulation"
author: "Beatriz Valcarcel"
date: "2022-04-20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Baseline data simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(ggplot2)
library(data.table)
library(magrittr)
library(surveillance)

```

## Simulate 16 baselines
Simulation of 16 different time series of count data. Parameter values for data simulation are presented in Noufaily et al. (2019), table 1.

### Each of the simulatation take into account a combination of differents properties of syndromic data such as: 

- Baseline frequencies of reports (determined by parameter **alpha**)
- Linear trends (determined by parameter**beta**)
- Seasonal trends (determined by parameters **gamma_1** and **gamma_2**), 
- Day-of-the week effects (determined by parameters **gamma_3** and **gamma_4**)
- Over dispersion (determined by parameter **phi**).



```{r}
baseline_1  <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 2,
                                    alpha = 6, 
                                    beta = 0, 
                                    gamma_1 = 0.2, 
                                    gamma_2 = 0.2, 
                                    gamma_3 = 0.5, 
                                    gamma_4 = 0.4, 
                                    phi = 2, 
                                    shift_1 = 29 
                                  )



baseline_1[,s:=1]

baseline_2  <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 2,
                                    alpha = 0.5, 
                                    beta = 0, 
                                    gamma_1 = 1.5, 
                                    gamma_2 = 1.4, 
                                    gamma_3 = 0.5, 
                                    gamma_4 = 0.4, 
                                    phi = 1, 
                                    shift_1 = -167 
                                  )


baseline_2[,s:=2]


baseline_3  <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 0,
                                    weekly_pattern_n = 2,
                                    alpha = 5.5, 
                                    beta = 0, 
                                    gamma_1 = 0, 
                                    gamma_2 = 0, 
                                    gamma_3 = 0.3, 
                                    gamma_4 = 0.25, 
                                    phi = 1, 
                                    shift_1 = 1
                                  )

baseline_3[,s:=3]


baseline_4 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 0,
                                    weekly_pattern_n = 2,
                                    alpha = 0, 
                                    beta = 0, 
                                    gamma_1 = 0, 
                                    gamma_2 = 0, 
                                    gamma_3 = 0.3,
                                    gamma_4 = 0.25, 
                                    phi = 1, 
                                    shift_1 = 1
                                  )

baseline_4[,s:=4]

baseline_5 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 2,
                                    alpha = 6, 
                                    beta = 0, 
                                    gamma_1 = 0.3, 
                                    gamma_2 = 2, 
                                    gamma_3 = 0.3, 
                                    gamma_4 = 0.5, 
                                    phi = 1.5, 
                                    shift_1 = -50 
                                  )

baseline_5[,s:=5]


baseline_6 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 1,
                                    alpha = 1, 
                                    beta = 0, 
                                    gamma_1 = 0.1, 
                                    gamma_2 = 2, 
                                    gamma_3 = 0.05, 
                                    gamma_4 = 0.05, 
                                    phi = 1, 
                                    shift_1 = -50 
                                  )

baseline_6[,s:=6]


baseline_7 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 0,
                                    weekly_pattern_n = 1,
                                    alpha = 6, 
                                    beta = 0.0001, 
                                    gamma_1 = 0, 
                                    gamma_2 = 0, 
                                    gamma_3 = 0.6, 
                                    gamma_4 = 0.9, 
                                    phi = 1.5, 
                                    shift_1 = 0 
                                  )


baseline_7[,s:=7]

baseline_8 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 1,
                                    alpha = 3, 
                                    beta = 0, 
                                    gamma_1 = 1.5,
                                    gamma_2 = 0.1, 
                                    gamma_3 = 0.2, 
                                    gamma_4 = 0.3, 
                                    phi = 1, 
                                    shift_1 = -150 )



baseline_8[,s:=8]

baseline_9 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 1,
                                    alpha = 3, 
                                    beta = 0, 
                                    gamma_1 = 0.2, 
                                    gamma_2 = 0.1, 
                                    gamma_3 = 0.05, 
                                    gamma_4 = 0.15, 
                                    phi = 1, 
                                    shift_1 = -200 
                                  )

baseline_9[,s:=9]


baseline_10 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 1,
                                    alpha = 5, 
                                    beta = 0, 
                                    gamma_1 = 0.2, 
                                    gamma_2 = 0.1, 
                                    gamma_3 = 0.05, 
                                    gamma_4 = 0.1, 
                                    phi = 1, 
                                    shift_1 = 0 
                                  )


baseline_10[,s:=10]


baseline_11 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 2,
                                    weekly_pattern_n = 1,
                                    alpha = 0.5, 
                                    beta = 0, 
                                    gamma_1 = 0.4, 
                                    gamma_2 = 0, 
                                    gamma_3 = 0.05, 
                                    gamma_4 = 0.15, 
                                    phi = 1, 
                                    shift_1 = 0 
                                  )

baseline_11[,s:=11]


baseline_12 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 2,
                                    weekly_pattern_n = 1,
                                    alpha = 9, 
                                    beta = 0, 
                                    gamma_1 = 0.5, 
                                    gamma_2 = 0.2, 
                                    gamma_3 = 0.2, 
                                    gamma_4 = 0.4, 
                                    phi = 4, 
                                    shift_1 = 57
                                  )


baseline_12[,s:=12]



baseline_13 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 2,
                                    weekly_pattern_n = 1,
                                    alpha = 2,
                                    beta = 0.0005, 
                                    gamma_1 = 0.8, 
                                    gamma_2 = 0.8, 
                                    gamma_3 = 0.8, 
                                    gamma_4 = 0.4, 
                                    phi = 4, 
                                    shift_1 = 57 
                                  )

baseline_13[,s:=13]



baseline_14 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 4,
                                    alpha = 0.05, 
                                    beta = 0, 
                                    gamma_1 = 0.01, 
                                    gamma_2 = 0.01, 
                                    gamma_3 = 1.8, 
                                    gamma_4 = 0.1, 
                                    phi = 1, 
                                    shift_1 = -85 
                                  )

baseline_14[,s:=14]


baseline_15 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 1,
                                    weekly_pattern_n = 2,
                                    alpha = 3, 
                                    beta = 0, 
                                    gamma_1 = 0.8, 
                                    gamma_2 = 0.6, 
                                    gamma_3 = 0.8, 
                                    gamma_4 = 0.4, 
                                    phi = 4, 
                                    shift_1 = 29)

baseline_15[,s:=15]



baseline_16 <- splalert::simulate_baseline_data(
                                    start_date = as.Date("2012-01-01"),
                                    end_date = as.Date("2019-12-31"),
                                    seasonal_pattern_n = 0,
                                    weekly_pattern_n = 2,
                                    alpha = 6, 
                                    beta = 0, 
                                    gamma_1 = 0, 
                                    gamma_2 = 0, 
                                    gamma_3 = 0.8, 
                                    gamma_4 = 0.4, 
                                    phi = 4, 
                                    shift_1 = 1
                                  )
                                   

baseline_16[,s:=16]

baseline <- rbind(baseline_1,
                  baseline_2,
                  baseline_3,
                  baseline_4,
                  baseline_5,
                  baseline_6,
                  baseline_7,
                  baseline_8,
                  baseline_9,
                  baseline_10,
                  baseline_11,
                  baseline_12,
                  baseline_13,
                  baseline_14,
                  baseline_15,
                  baseline_16)

baseline[, label:=paste("Sim", s, sep=" ")]

q <- ggplot(baseline, aes(x = date, y = n))
q <- q + geom_line(lwd = 1)
q <- q + facet_wrap(~s, scales = "free")
q <- q + splstyle::scale_fill_fhi(palette = "contrast")
q <- q + splstyle::theme_fhi_lines_horizontal(legend_position = "bottom")
q
```

## Weekly patterns

```{r}
q <- ggplot(baseline, aes(wday, n, group=wday))
q <- q + facet_wrap(~s, scales = "free")
q <- q + geom_boxplot()
q
```

# Monthly patterns

```{r}
q <- ggplot(baseline, aes(wday, n, group=calmonth))
q <- q + facet_wrap(~s, scales = "free")
q <- q + geom_boxplot()
q

```

