---
title: "Short term trend for COVID cases Part 2: map"
author: "Chi Zhang"
date: "2022-05-30"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Short term trend for COVID cases Part 2: map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we illustrate how to compute short term trend with `splalert::short_term_trend` for multiple locations, and visualize in a map. We also include labels with the number of confirmed COVID cases per 100 000 population (for each county).

```{r}
library(ggplot2)
library(data.table)
library(magrittr)
```

## Covid-19 MSIS cases

The Covid-19 cases (lab confirmed) for Norway, for each county and the whole nation. The data does not distinguish age groups or sex.

This dataset is extracted on 2022-05-04. The time period is from 2020-02-21 to 2022-05-03.

The data in csv and xlsx formats can be found on our [Github repository](https://github.com/folkehelseinstituttet/surveillance_data).

### Data in `spltidy` format

We have prepared the data into [spltidy format](https://docs.sykdomspulsen.no/spltidy/articles/splfmt_rts_data_v1.html), from which we can see the summary of each column, such as data type and amount of missing data.

```{r}
d_msis <- spltidy::covid19_msis_cases_by_time_location
# check the column names
# colnames(d_msis)  
```

## 1. Plan analysis for all locations with `plnr`

First define a new plan, then add data in the following way.

```{r}
# We begin by defining a new plan
p <- plnr::Plan$new()

# 1. add data 
data_fn <- function(){
  return(d_msis)
}

p$add_data(
  name = "covid19_cases",
  fn_name = "data_fn"
)

p$get_data() %>% names
```

Next we add some argument sets (we call them *argset*). We want to repeat the same analysis for all locations available in this dataset, which are 1 nation and 11 counties. It can be done by `expand_list()`.

```{r}
# 2. add argset 
# check location codes
location_codes <- p$get_data()$covid19_cases$location_code %>%
  unique() %>%
  print()

p$add_argset_from_list(
  plnr::expand_list(
    location_code = location_codes,
    granularity_time = "isoweek"
  )
)
# Examine the argsets that are available
p$get_argsets_as_dt()

```

Next we create an **action**. This allows us to develop and interact with only one data specification (e.g. `location_code == 'county03'`) without having to do explicit filtering.

On this specific subset, we carry out one function like in the previous example, `splalert::shorterm_trend`. Return the result at the end of the action function.

```{r}
# 3. action
# To do this, we first need to create an action function
# (takes two arguments -- data and argset)

action_fn <- function(data, argset){
  if(plnr::is_run_directly()){
    data <- p$get_data()
    argset <- p$get_argset(1)  # county03, isoweek
  }
  # develop function for ONE argset only

  # data
  pd <- data$covid19_cases[
    location_code == argset$location_code &
      granularity_time == argset$granularity_time
  ]

  # function
  trend_msis <- splalert::short_term_trend(
    pd,
    numerator = "covid19_cases_testdate_n",
    trend_isoweeks = 6,
    remove_last_isoweeks = 1
  )
  # return
  trend_msis

}
```

Now we apply the action function on all 12 data specifications. 

```{r}
# apply this to all 12 argsets
p$apply_action_fn_to_all_argsets(fn_name = "action_fn")
```


## 2. Run analysis for all 12 argsets 

You can run the analysis one by one, or run all 12 together.

```{r}
# run one by one
# p$run_one(1)
# p$run_one(2)

# run together
res <- p$run_all_progress()
```

The results are saved in a list of length 12. You can access them in the same way you access list. 

```{r}
res_county03 <- res[[1]]
print(res_county03[, .(location_code, date, covid19_cases_testdate_n, covid19_cases_testdate_trend0_42_status)])
```

We can collect the results from each list into one data table.

```{r}
res_unlisted <- rbindlist(res)
```


## 3. visualize in a map 

```{r}
# select data for one week (narrow data)
this_isoyearweek <- '2021-44'
d_msis_this_isoyearweek <- res_unlisted[granularity_geo == 'county' &
                                   isoyearweek == this_isoyearweek,
                                   .(location_code,
                                     date,
                                     isoyearweek,
                                     covid19_cases_testdate_pr100000,
                                     covid19_cases_testdate_trend0_42_status)]


```


Now we make a map 

```{r}
pd <- copy(splmaps::norway_nuts3_map_b2020_insert_oslo_dt)

d_msis_this_isoyearweek[, trend := factor(
  covid19_cases_testdate_trend0_42_status,
  levels = c('decreasing','null', 'increasing')
)]



# join the map data.table
pd[d_msis_this_isoyearweek,on="location_code",trend:=trend]

# plot map
q <- ggplot()
q <- q + geom_polygon(
  data = pd,
  mapping = aes(x = long, y = lat, group = group,fill=trend),
  color="black",
  size=0.25
)
q <- q + coord_quickmap()
q <- q + theme_void()
q <- q + labs(title=glue::glue("MSIS cases per 100k population for week ", this_isoyearweek))
q <- q + splstyle::scale_fill_fhi("Category",palette = "map_seq_complete", direction = 1, drop=F)
q <- q + labs(caption = "Data updated 2021 week 44. Demonstration only")

# add watermark later!
q
```

Add label 

```{r}
# now we add label to the map
# which contains county name + case per 100k
labels <- copy(splmaps::norway_nuts3_position_geolabels_b2020_insert_oslo_dt)
labels[
  d_msis_this_isoyearweek,
  on = "location_code",
  cases_100k := covid19_cases_testdate_pr100000
]

labels[
  spldata::norway_locations_names_b2020,
  on = "location_code",
  location_name := location_name
]

# format case, remove decimal
labels[, cases_100k := splstyle::format_nor_num_0(cases_100k)]
labels

# put case 100k together with county name
labels[, label := paste0(location_name, '\n', cases_100k)]

q <- q + ggrepel::geom_label_repel(
  data = labels,
  mapping = aes(x = long, y = lat, label = label)
)
q
```

