---
title: "Short term trend"
author: "Richard Aubrey White, Chi Zhang"
date: "2022-05-05"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Short term trend}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


In this vignette we illustrate how to compute short term trend with `splalert::short_term_trend`. The data we use is the Covid-19 hospitalization in Norway. 


```{r}
library(ggplot2)
library(data.table)
library(magrittr)
```

## Covid-19 Hospitalization data

The Covid-19 hospitalization data contains both daily and weekly aggregated number of hospitalization due to Covid (as a main cause) for Norway. The data does not distinguish age groups or sex. 

This dataset is extracted on 2022-05-04. The time period is from 2020-02-21 to 2022-05-03. 

The data in csv and xlsx formats can be found on our [Github repository](https://github.com/folkehelseinstituttet/surveillance_data). 


### Data in `spltidy` format

We have prepared the data into [spltidy format](https://docs.sykdomspulsen.no/spltidy/articles/splfmt_rts_data_v1.html), from which we can see the summary of each column, such as data type and amount of missing data. 

```{r}
d_hosp <- spltidy::norway_covid19_icu_and_hospitalization
# check the column names
colnames(d_hosp)  

d_hosp
```









## Weekly observations

Now we run the `short_term_trend` function on weekly data. 

```{r}
d_hosp_weekly <- d_hosp[granularity_time=="isoweek"]


res <- splalert::short_term_trend(
  d_hosp_weekly, 
  numerator = "hospitalization_with_covid19_as_primary_cause_n",
  trend_isoweeks = 6,
  remove_last_isoweeks = 1
)

# create the trend label
res[, hospitalization_with_covid19_as_primary_cause_trend0_42_status := factor(
  hospitalization_with_covid19_as_primary_cause_trend0_42_status,
  levels = c("training","forecast","decreasing", "null", "increasing"),
  labels = c("Training","Forecast","Decreasing", "Null", "Increasing")
)]

colnames(res)
```


We can check some columns that have been added to the original data. 
```{r}
# check some columns 
res[, .(date, 
        hospitalization_with_covid19_as_primary_cause_n, 
        hospitalization_with_covid19_as_primary_cause_forecasted_n,         hospitalization_with_covid19_as_primary_cause_trend0_42_status)]
```

We can visualize the trend indicator with different colors. We use [splstyle](https://docs.sykdomspulsen.no/splstyle/) customized ggplot2 syntax for standardized graphic styles. 


```{r}
q <- ggplot(res, aes(x = isoyearweek, y = hospitalization_with_covid19_as_primary_cause_forecasted_n, group = 1))
q <- q + geom_col(mapping = aes(fill = hospitalization_with_covid19_as_primary_cause_trend0_42_status ))
q <- q + geom_errorbar(
  mapping = aes(
    ymin = hospitalization_with_covid19_as_primary_cause_forecasted_predinterval_q02x5_n,
    ymax = hospitalization_with_covid19_as_primary_cause_forecasted_predinterval_q97x5_n
  )
)
q <- q + scale_y_continuous("Weekly hospitalization with Covid-19 as primary cause", expand = c(0, 0.1))
q <- q + scale_x_discrete("Isoyearweek", breaks = splstyle::every_nth(8))
q <- q + expand_limits(y=0)
q <- q + splstyle::scale_fill_fhi("6 week trend", palette = "contrast")
q <- q + splstyle::theme_fhi_lines_horizontal(legend_position = "bottom")
q <- q + splstyle::set_x_axis_vertical()
q
```






## Daily observations

Similarly, we can also do the analysis using daily data.

```{r}
d_hosp_daily <- d_hosp[granularity_time=="day"]

res <- splalert::short_term_trend(
  d_hosp_daily, 
  numerator = "hospitalization_with_covid19_as_primary_cause_n",
  trend_days = 28,
  remove_last_days = 7
)
res[, hospitalization_with_covid19_as_primary_cause_trend0_28_status := factor(
  hospitalization_with_covid19_as_primary_cause_trend0_28_status,
  levels = c("training","forecast","decreasing", "null", "increasing"),
  labels = c("Training","Forecast","Decreasing", "Null", "Increasing")
)]

```


We can also visualize this trend figure.
```{r}
q <- ggplot(data = res, mapping = aes(x = date, y = hospitalization_with_covid19_as_primary_cause_forecasted_n, fill = hospitalization_with_covid19_as_primary_cause_trend0_28_status))
q <- q + geom_col(alpha = 0) # necessary to get legend to be in right order
q <- q + geom_col(
  data = res[hospitalization_with_covid19_as_primary_cause_trend0_28_status!="Forecast"]
  )
q <- q + geom_ribbon(
  data = res[hospitalization_with_covid19_as_primary_cause_trend0_28_status=="Forecast"], 
  mapping = aes(
    ymin = hospitalization_with_covid19_as_primary_cause_forecasted_predinterval_q02x5_n,
    ymax = hospitalization_with_covid19_as_primary_cause_forecasted_predinterval_q97x5_n
  ),
  alpha = 0.75
)
q <- q + scale_y_continuous("Daily hospitalization with Covid-19 as primary cause", expand = c(0, 0.1))
q <- q + scale_x_date("Date", date_breaks = "28 days", date_labels = "%y-%m-%d")
# q <- q + scale_x_discrete(breaks = splstyle::every_nth(8))
q <- q + expand_limits(y=0)

q <- q + splstyle::scale_fill_fhi("28 days trend", palette = "contrast")
q <- q + splstyle::theme_fhi_lines_horizontal(legend_position = "bottom")
q <- q + splstyle::set_x_axis_vertical()
q
```


Another way to visualize trend

```{r}
q <- ggplot(res, aes(x = date, y = hospitalization_with_covid19_as_primary_cause_doublingdays0_28))
q <- q + geom_rect(aes(xmin = date-1, xmax=date, ymin = 1, ymax = Inf, fill = hospitalization_with_covid19_as_primary_cause_trend0_28_status), alpha = 0.5)
q <- q + geom_line(lwd = 1)
q <- q + scale_y_continuous(trans = "log10", expand = c(0, 0.1))
q <- q + splstyle::scale_fill_fhi(palette = "contrast")
q <- q + splstyle::theme_fhi_lines_horizontal(legend_position = "bottom")
q

```
