---
title: "Short term trend for COVID cases Part 1 epicurve"
author: "Chi Zhang"
date: "2022-05-30"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Short term trend for COVID cases Part 1 epicurve}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we illustrate how to compute short term trend with `splalert::short_term_trend`.

```{r}
library(ggplot2)
library(data.table)
library(magrittr)
```

## Covid-19 MSIS cases

The Covid-19 cases (lab confirmed) for Norway, for each county and the whole nation. The data does not distinguish age groups or sex.

This dataset is extracted on 2022-05-04. The time period is from 2020-02-21 to 2022-05-03.

The data in csv and xlsx formats can be found on our [Github repository](https://github.com/folkehelseinstituttet/surveillance_data).

### Data in `spltidy` format

We have prepared the data into [spltidy format](https://docs.sykdomspulsen.no/spltidy/articles/splfmt_rts_data_v1.html), from which we can see the summary of each column, such as data type and amount of missing data.

```{r}
d_msis <- spltidy::norway_covid19_cases_by_time_location
# check the column names
colnames(d_msis)  
```

### Understand the data

(granularity geo and location codes)

(one visualization of counts for each county)

```{r}
d_fylke <- copy(d_msis[granularity_time == 'isoweek'& granularity_geo == 'county'] )
q <- ggplot(d_fylke, aes(x = isoyearweek, y = covid19_cases_testdate_n))
q <- q + geom_line(aes(group = location_code, col = location_code),size = 2)
q <- q + scale_y_continuous("Weekly MSIS cases",
                            expand = expansion(mult = c(0, 0.1)),
                            labels = splstyle::format_nor_num_0)

q <- q + scale_x_discrete("Isoyearweek",
                          breaks = splstyle::every_nth(8))

q <- q + splstyle::scale_color_fhi("Location", palette = "primary")
q <- q + splstyle::theme_fhi_lines_horizontal(legend_position = "bottom")
q <- q + splstyle::set_x_axis_vertical()
# need watermark!!
q

```

(another visualization of pr, including the nation)

```{r}
q <- ggplot(d_fylke, aes(x = isoyearweek, y = covid19_cases_testdate_pr100000))
q <- q + geom_line(aes(group = location_code, col = location_code),size = 2)
q <- q + scale_y_continuous("Weekly MSIS cases (per 100 000)",
                            expand = expansion(mult = c(0, 0.1)),
                            labels = splstyle::format_nor_num_0)

q <- q + scale_x_discrete("Isoyearweek",
                          breaks = splstyle::every_nth(8))

q <- q + splstyle::scale_color_fhi("Location", palette = "primary")
q <- q + splstyle::theme_fhi_lines_horizontal(legend_position = "bottom")
q <- q + splstyle::set_x_axis_vertical()
# need watermark!!
q

```

## 1. weekly trend

(maybe not the whole period, only visualise one year)

(nation, each county)

```{r}
d_msis_norge <- copy(d_msis[granularity_time == 'isoweek' &
                         isoyearweek >= '2021-26' &
                         location_code == 'norge'])

trend_msis_norge <- splalert::short_term_trend(
  d_msis_norge,
  numerator = "covid19_cases_testdate_n",
  trend_isoweeks = 6,
  remove_last_isoweeks = 1
)


# colnames(trend_msis_norge)

# attach status for vis
trend_msis_norge[, covid19_cases_testdate_trend0_42_status := factor(
  covid19_cases_testdate_trend0_42_status,
  levels = c("training","forecast","decreasing", "null", "increasing"),
  labels = c("Training","Forecast","Decreasing", "Null", "Increasing")
)]


# plot
q <- ggplot(trend_msis_norge, aes(x = isoyearweek,
                            y = covid19_cases_testdate_forecasted_n,
                            group = 1))
q <- q + geom_col(mapping = aes(fill = covid19_cases_testdate_trend0_42_status ))
q <- q + geom_errorbar(
  mapping = aes(
    ymin = covid19_cases_testdate_forecasted_predinterval_q02x5_n,
    ymax = covid19_cases_testdate_forecasted_predinterval_q97x5_n
  )
)
q <- q + scale_y_continuous("Weekly MSIS cases",
                            expand = c(0, 0.1),
                            labels = splstyle::format_nor_num_0)
q <- q + scale_x_discrete("Isoyearweek", breaks = splstyle::every_nth(6))
q <- q + expand_limits(y=0)
q <- q + splstyle::scale_fill_fhi("6 week trend", palette = "contrast")
q <- q + splstyle::theme_fhi_lines_horizontal(legend_position = "bottom")
q <- q + splstyle::set_x_axis_vertical()
q

```

In the part 2 of the COVID analysis, we demonstrate how to do the trend analysis for all the counties, and visualize the results in a map. 



