---
title: "Short term trend: MSIS cases"
author: "Chi Zhang"
date: "2022-05-30"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Short term trend: MSIS cases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


In this vignette we illustrate how to compute short term trend with `splalert::short_term_trend`.


```{r}
library(ggplot2)
library(data.table)
library(magrittr)
```

## Covid-19 MSIS cases

The Covid-19 XXXXXXXXXXX  for Norway. The data does not distinguish age groups or sex. 

This dataset is extracted on 2022-05-04. The time period is from 2020-02-21 to 2022-05-03. 

The data in csv and xlsx formats can be found on our [Github repository](https://github.com/folkehelseinstituttet/surveillance_data). 


### Data in `spltidy` format

We have prepared the data into [spltidy format](https://docs.sykdomspulsen.no/spltidy/articles/splfmt_rts_data_v1.html), from which we can see the summary of each column, such as data type and amount of missing data. 

```{r}
d_msis <- spltidy::covid19_msis_cases_by_time_location
# check the column names
colnames(d_msis)  
```



### Understand the data

(granularity geo and location codes)


(one visualization of counts for each county)

(another visualization of pr, including the nation)






## 1. weekly trend 
(maybe not the whole period, only visualise one year)

(nation, each county)





## 2. Trend of last week in map 






